### 2025-03-31 chapter/paper.tex
small revisions to chapter

\documentclass[]{article}
\usepackage[round]{natbib}

\usepackage{fullpage}
\usepackage{url}
\usepackage{authblk}
\usepackage{graphicx}
\usepackage{color}

% need to have the pythonhighlight.sty in directory
%\usepackage{pythonhighlight}

% use minted for python and YAML code
\usepackage{minted}

% local definitions
\newcommand{\aprcomment}[1]{{\textcolor{blue}{APR: #1}}}
\usepackage{xspace}
\newcommand{\moments}{\texttt{moments}\xspace}
\newcommand{\demes}{\texttt{demes}\xspace}
\newcommand{\demesdraw}{\texttt{demesdraw}\xspace}
\newcommand{\msprime}{\texttt{msprime}\xspace}
\newcommand{\momi}{\texttt{momi}\xspace}
\newcommand{\fwdpy}{\texttt{fwdpy11}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

\begin{document}

\title{Demographic inference using the SFS with \moments and \demes}
\author[1]{Nick Collier}
%\author[2]{Simon Gravel(?)}
\author[1,*]{Aaron P. Ragsdale}
\affil[1]{Department of Integrative Biology, University of Wisconsin--Madison}
%\affil[2]{McGill University}
\affil[*]{apragsdale@wisc.edu}
\maketitle

\begin{abstract}
In this chapter we show how to use \moments, a Python library for computing the
joint SFS and inferring demographic history, through its \demes interface. 
Following a discussion of SFS inference, we illustrate the general usage of 
\moments and sketch two specific examples. We conclude with a discussion of some 
additional considerations regarding the \moments library.
\end{abstract}

\section*{Introduction}

The genetic composition of a sample of individuals is shaped by their genome
biology and evolutionary history. Variation resulting from this history can be
fully represented by the ancestral relationships among samples at each locus in
the genome and how those gene-genealogical relationships change along a
chromosome due to recombination (that is, by information stored in the Ancestral
Recombination Graph \citep{nielsen2025inference}). However, the ARG can be
large and unwieldy, and methods for reconstructing history directly from the
ARG, while showing promise \citep[e.g.,][]{yc2022evaluation, fan2023likelihood,
brandt2024promise}, are in their infancy and so far limited in application and
scalability. Instead, evolutionary inference using informative summaries of
genetic variation remains a tractable and powerful alternative for learning
parameters of population history, natural selection and genome biology.

One such summary that has seen wide use is the site frequency spectrum (SFS),
which stores the counts (observed or expected) of alleles carried by a given
number of genomes in a set of samples. Like any summary of the data, the SFS
discards some information stored in the ARG -- in this case, loci are treated
independently so that haplotypic information is lost. Even so, the relative
densities of allele frequencies and the overall scale of the SFS are both
sensitive to demographic and non-demographic processes.

This chapter focuses on multi-population demographic inference from the SFS
using \moments \citep{jouganous2017inferring}.  Past population processes,
including changes in population size, splits, gene flow and population
structure, impact the SFS in predictable ways. A population size expansion (or
contraction) leads to a relative excess (or deficit) of low-frequency variants,
for example, and commonly used statistics measuring population divergence, such
as $F_{ST}$, are themselves summaries of the SFS.  Thus, the development of
methods to learn demographic history from the SFS came soon after the
publication of whole-genome sequencing from multiple individuals
\citep{marth2004allele, williamson2005simultaneous}. More recent methodological
advances allow for increased sample sizes and numbers of populations, providing
a rich ecosystem of software for SFS-based demographic inference
\citep{gutenkunst2009inferring, excoffier2011fastsimcoal,
gravel2011demographic, jouganous2017inferring, ragsdale2018genomic,
kamm2020efficiently, dilber2024faster}. We return to this in the final section:
Considerations and Caveats.

Other evolutionary mechanisms are known to affect allele frequency dynamics and
the SFS. This presents a challenge for demographic inference, as the observed
SFS may be distorted by non-demographic processes. However, this also presents
an opportunity to learn about different evolutionary processes, if demographic
history can be controlled for. The SFS has been used to infer the distribution
of fitness effects of new mutations, primarily for nonsynonymous variation in
coding regions \citep[e.g.,][]{eyre2006distribution, boyko2008assessing,
kim2017inference}; to scan for historical selective sweeps
\citep[e.g.,][]{kim2002detecting, nielsen2005genomic}; and to understand how
selection on quantitative traits impacts the genetic architecture underlying
those traits \citep[eg.,][]{patel2024conditional, ragsdale2024archaic}. The SFS
has also been used to infer relative mutation rates and how they have changed
over time \citep{dewitt2021nonparametric}. Each of these analyses typically
requires partitioning the data in some way -- by functional annotation,
mutation context, local recombination rate, etc -- so that comparisons can be
made across SFS from different classes of mutations.

\subsection*{Interfacing \moments with \demes}

Specifying demographic models requires defining populations (or ``demes'') and
their relationships via splits and gene flow. \demes provides a standardized
and accessible format for defining demographic models \citep{gower2022demes},
and has been adopted by widely used software for population genetics simulation
and inference (including \msprime \citep{baumdicker2022efficient}, \momi
\citep{dilber2024faster}, \fwdpy \citep{thornton2019polygenic}, \texttt{GADMA}
\citep{noskova2023gadma2}). \demes encourages interoperability and reuse of
code across software, ease of model implementation and reduction of errors.

Demographic models are specified in YAML format (see \citet{gower2022demes} and
the associated documentation). The two-population split-with-migration model
depicted in Figure~\ref{fig:im}A is specified as below:
\begin{minted}[gobble=4, frame=single,]{yaml}
    description: split-with-migration model, loosely based on example 2
    generation_time: 29
    time_units: years
    demes:
    - name: ancestral
      epochs:
      - {start_size: 15000, end_time: 300000}
      - {start_size: 25000, end_time: 75000}
    - name: popA
      ancestors: [ancestral]
      epochs:
      - {start_size: 30000}
    - name: popB
      ancestors: [ancestral]
      epochs:
      - {start_size: 1200, end_size: 14000}
    migrations:
    - demes: [popA, popB]
      rate: 5e-5
\end{minted}

In \moments, a \demes-specified demographic model can be directly used to
compute either the SFS or multi-population LD statistics
\citep{ragsdale2019models, ragsdale2020unbiased}. For the SFS, we simply need
to load the demographic model using \demes and specify the number of haploid
samples to draw and from which populations.
\begin{minted}[gobble=4, frame=single]{python}
    import demes, moments
    g = demes.load('model.yaml')
    samples = {'popA': 60, 'popB': 60}
    fs = moments.Demes.SFS(g, samples=samples)
\end{minted}
Here, we have not specified a mutation rate. The default behavior sets $u=1$,
so that $\theta = 4N_eu=4N_e$, where $N_e$ is the ancestral population's
initial size. With a known mutation rate, the SFS will be properly scaled by
passing the mutation rate as a keyword argument: \texttt{moments.Demes.SFS(g,
samples=samples, u=u)}. This is equivalent to scaling the output SFS above by
multiplying by \texttt{u}, as \texttt{u*fs}.

\section*{Examples: inference in \moments using \demes}

Below, we show two examples of multi-population demographic inference using the
joint SFS. The first example simulates data with known demographic parameters,
which we try to reinfer using the both the original model and simpler misspecified 
ones. In the second example, we infer a relatively simple human-Neanderthal
history using data from two human populations and a high-coverage ancient
sample from the Neanderthal lineage. For both of these examples, we briefly
describe data processing and inference setup, emphasizing the high-level
components of the analyses. For detailed information about each example,
including managing data, specifying models, performing inference, computing
confidence intervals and visualizing fits, we refer readers to the accompanying
GitHub repository
(\url{https://github.com/StatisticalPopulationGenomics-2ndEd/moments}), which
is maintained with up-to-date versioning.

\begin{figure}[t!]
    \includegraphics{../example1/fig1.pdf}
    \caption{Caption placeholder}
    \label{fig:im}
\end{figure}

In order to optimize parameters in a \demes model, we need a way to specify
which parameters should be fit. For this, we use a separate YAML-formatted file
to define \texttt{parameters}, each of which points to desired value(s) in the
demographic model and sets lower and upper bounds. This file also specifies any
relative \texttt{constraints} between pairs of parameters, to ensure the
optimization routine only explores valid model space. Below, we have included
a truncated parameters file -- the full example is available on GitHub.
\begin{minted}[gobble=4, frame=single]{YAML}
    parameters:
    - name: T0
      values:
      - demes:
          ancestral:
            epochs:
              0: end_time
      lower_bound: 0
      upper_bound: 2e5
    - name: T1
      values:
      - demes:
          ancestral:
            epochs:
              1: end_time
      lower_bound: 0
      upper_bound: 1e6
    - name: Ne
      values:
      - demes:
          ancestral:
            epochs:
              0: start_size
      lower_bound: 1e2
      upper_bound: 1e6
    ...
    constraints:
    - params: [T0, T1]
      constraint: greater_than
\end{minted}

Taking the demographic model, parameters file, and observed SFS together, we
can perform inference using the \texttt{moments.Demes.Inference.optimize(...)}
function.  The initial parameter guesses are given by the input demographic
model, which may be perturbed using a keyword argument. Any value in the
demographic model that is not specified in the parameters file will remain
unchanged, and is therefore treated as a fixed parameter. Here, we also assume
we have an estimate for the total mutation rate, $U$.
\begin{minted}[gobble=4, frame=single]{python}
    import moments

    data = moments.Spectrum.from_file('data.fs')
    U = 1e-8 * 5e8  # per-base mutation rate times the total length

    model_file = 'model.yaml'
    params_file = 'params.yaml'
    output = 'model.fit.yaml'

    ret = moments.Demes.Inference.optimize(
        model_file, params_file, data,
        perturb=1, uL=U, output=output
    )
\end{minted}

Confidence intervals may also be computed using 
\texttt{moments.Demes.Inference.uncerts(...)}, once a (locally) optimal model 
has been found. The following will emit a tab-separated file containing
parameter names, best-fit values and estimated standard errors.
\begin{minted}[gobble=4, frame=single]{python}
    log_file = 'log.txt'

    uncerts = moments.Demes.Inference.uncerts(
      output, params_file, data,
      uL=U, output=log_file
    )
\end{minted}

\subsection*{Inferring parameters in a simulated split-with-migration model}

Using the split-with-migration model defined above (illustrated in
Figure~\ref{fig:im}A), we simulated data for 30 diploid individuals from both
populations using \msprime \citep{baumdicker2022efficient}. These simulations
consided of 500 regions, each of length 1 Mb, with per-base recombination and
mutation rates of $10^{-8}$. The joint SFS was computed using \tskit's
\texttt{ts.allele\_frequency\_spectrum()} function separately on each replicate
region and summing across regions to find the total SFS across 500 Mb of data
(Figure~\ref{fig:im}B--D).

We fit three demographic models to the simulated data. In addition to
reinferring the parameters from the simulated model, we fit two simpler models
to the same data. These had fewer parameters, both omitting the size change
deeper in time in the ancestral population. This is meant to crudely mimic the
scenario that we often face, in which the true history is more complicated than
the parameterized model. It also highlights biases in inferences that can arise
when features of the true history are not included.

When fitting the two misspecified models that do not allow for size changes in
the ancestral population, this split time is inferred to be substantially
deeper in the past than the true split time. This effect, as demonstrated by
\citet{momigliano2020biases}, is due to the decreased coalescence rate within
the ancestral population between the time of the expansion and divergence of
the descendent populations. These two misspecified models
(Figure~\ref{fig:im}H,I) also provide a worse fit to the data, as seen by the
lower log-likelihoods and the large residuals between model predictions and
data (Figure~\ref{fig:im}E-G).

\subsection*{Inferring human-Neanderthal demographic parameters}

We obtained modern human genome sequences from a recent resequencing of the 
1000 Genomes Project cohort \citep{byrska2022high}, and used the high-coverage 
sequence of the Vindija Neandertal from \cite{prufer2017high}. 
From the 1000 Genomes cohort we took 85 MSL (Mende from Sierre Leone) and 91 
GBR (British from England and Scotland) sequences as our modern human samples.  
We filtered out sites that lay outside the 1000 Genomes `strict' mask, 
were not genotyped in all four existing high-coverage archaic genome sequences, 
thatlacked high-confidence ancestral state assignments, or fell within 10 kb of 
exonic or promoter regions and were left with approximately 960 Mb of 
well-characterized and putatively neutrally-evolving sequence.
We estimated the genome-wide SFS using built-in \moments parsing functions.
Details concerning data processing (including our liftover of the Vindija 
sequence to a more recent genome build) can be found in the linked GitHub 
repository.

The model that we wished to fit is relatively simple but still has many 
parameters (11 in the example we present), so we constructed it in stages by 
first fitting simpler one-- and two--population models.
Throughout, we used $u=1.5\cdot 10^{-8}$ as an estimate of the mutation rate per 
generation and 29 years as the average generation time. We began by fitting 
one-population models with ancestral and recent size changes for MSL and GBR.
For the MSL model, we allowed three epochs with constant effective size in each 
epoch. We attributed to GBR a sharp contraction $\sim$60 kya followed by 
exponential growth to represent the out-of-Africa bottleneck and subsequent 
rapid expansion. YAML files representing the models and parameters in this 
section can be found in the linked GitHub repository.

We next stitched these one-population models together and refitted the 
resulting graph to get a two-population model that included continuous symmetric 
migration between MSL and GBR and an ancestral size expansion. Following this,
we added a Neandertal branch and fitted the 
Neandertal-modern human divergence time, Neandertal effective size, and the 
Neandertal-to-GBR admixture proportion. We also fitted the divergence time 
between the sampled Vindija Neandertal lineage and the Neandertal lineage which 
admixed with modern non-African humans. Finding that this parameter tended to 
converge to its lower bound (imposed by the estimated time of death of the 
Vindija Neandertal, ~55 kya), we fixed it to a value supported by literature 
(90 kya, \cite{prufer2017high}) in a second round of optimization.
Rather than fitting it, we fixed the time of the Neandertal-GBR pulse to 50 kya.
Fixing parameters for which good estimates are available in the literature 
(see e.g. \cite{sumer2025earliest}) may be reasonable in the first stages of 
model optimization, and we expected this value to be especially difficult to fit 
due to the small pulse proportion and Neandertal sample size. In the second 
round of optimization on the three-population model, we refit all parameters 
aside from the two event times mentioned above to refine the estimates.

After this refitting step, we were left with parameters in good agreement with 
prior authors, with a Neandertal-modern human divergence time of 547 kya, 
MSL-GBR split time of 77 kya, and Neandertal-GBR pulse proportion of 2.7 
percent. We observe some large residuals in our best-fit model 
Figure~\ref{fig:humans}G-I), suggesting ...

\begin{figure}[t!]
    \includegraphics{../example2/fig2.pdf}
    \caption{
      A plot of the best-fit model created using \demesdraw is displayed in (A).
      Marginal SFS fits between the model and empirical data are shown for MSL 
      (B) and GBR (C).
      Marginal two-way spectra predicted by the model and their residuals 
      against the data are plotted for the three pairs of populations MSL, GBR 
      (D and G), MSL, Vindija (E and H), and GBR, Vindija (F, I).
    }
    \label{fig:humans}
\end{figure}

\section*{Considerations and caveats}\label{sec:conclusions}

\begin{enumerate}
    \item Include here general ideas about the strengths and weaknesses of
        various approaches in population genetic inference, including when
        using the SFS.
    \item Challenges in finding local/global optima.
    \item Challenges in exploring parameter space.
    \item Background selection [Ewing, Johri]
    \item Gene dense vs gene sparse genomic architectures among species.
\end{enumerate}

\bibliographystyle{plainnat}
\bibliography{paper}

\end{document}