### 2025-08-26 chapter/paper.tex
final version of revision 1

\documentclass[]{article}
\usepackage[numbers,sort,compress]{natbib}

\usepackage{url}
\usepackage{graphicx}
\usepackage{color}
\usepackage{setspace}
\usepackage{lineno}

\usepackage{authblk}
\title{Demographic inference using the SFS with \moments and \demes}
\author[1]{Nicholas W. Collier}
\author[1,*]{Aaron P. Ragsdale}
\affil[1]{Department of Integrative Biology, University of Wisconsin--Madison}
\affil[*]{apragsdale@wisc.edu}
\affil[ ]{ORCID IDs: 0009-0005-0385-9798 (NWC), 0000-0003-0715-3432 (APR)}

\usepackage[margin=1in, headheight=15pt, headsep=10pt]{geometry}
\usepackage{fancyhdr}
\pagestyle{fancy}
\lhead{NW Collier and AP Ragsdale}
\rhead{Demographic inference with \moments and \demes}

% use minted for python and YAML code
\usepackage{minted}

% local definitions
\usepackage{xspace}
\newcommand{\aprcomment}[1]{{\textcolor{blue}{APR: #1}}}
\newcommand{\reviewercomment}[1]{{\textcolor{red}{REVIEWER: #1}}}
\newcommand{\nwccomment}[1]{{\textcolor{magenta}{NWC: #1}}}
\newcommand{\moments}{\texttt{moments}\xspace}
\newcommand{\demes}{\texttt{demes}\xspace}
\newcommand{\demesdraw}{\texttt{demesdraw}\xspace}
\newcommand{\msprime}{\texttt{msprime}\xspace}
\newcommand{\momi}{\texttt{momi}\xspace}
\newcommand{\fwdpy}{\texttt{fwdpy11}\xspace}
\newcommand{\tskit}{\texttt{tskit}\xspace}

\begin{document}
\linenumbers
\doublespacing

\maketitle

\begin{abstract}

    The site frequency spectrum (SFS) describes the distribution of allele
    frequencies in one or more populations and sees wide use in evolutionary
    inference. Here we show how to use \moments, a Python library for working
    with the SFS and other measures of genetic diversity, and \demes, a
    standard format for parameterizing population models, to infer demographic
    history. Following an introduction to SFS inference, we illustrate the
    general usage of \moments and its \demes interface, and then show two
    specific examples, using both simulated and human genetic data. We conclude
    with caveats and considerations regarding the \moments library and SFS
    inference more broadly.

\end{abstract}

\emph{Keywords:} Demographic inference, site frequency spectrum, diffusion
equation

\section*{Introduction}

The genetic composition of a sample of individuals is shaped by their genome
biology and evolutionary history. Variation resulting from this history can be
fully represented by the ancestral relationships among samples at each locus in
the genome and how those gene-genealogical relationships change along a
chromosome due to recombination (that is, by information stored in the
Ancestral Recombination Graph; see chapters 1, 8 and 11, and
\cite{nielsen2025inference}). However, the ARG can be large and unwieldy, and
methods for reconstructing history directly from the ARG, while showing promise
(e.g., \cite{yc2022evaluation, fan2023likelihood, brandt2024promise}), are in
their infancy and so far limited in application and scalability. Instead,
evolutionary inference using informative summaries of genetic variation remains
a tractable and powerful alternative for learning parameters of population
history, natural selection and genome biology.

One such summary that has seen wide use is the site frequency spectrum (SFS),
which stores the counts (observed or expected) of alleles carried by a given
number of genomes in a set of samples. Like any summary of the data, the SFS
discards some information stored in the ARG -- in this case, loci are treated
independently so that haplotypic information is lost. Even so, the relative
densities of allele frequencies and the overall scale of the SFS are both
sensitive to demographic and non-demographic processes.

Past population processes, including changes in population size, splits, gene flow
and population structure, impact the SFS in predictable ways. A population size
expansion (or contraction) leads to a relative excess (or deficit) of
low-frequency variants, for example, and commonly used statistics measuring
population divergence, such as $F_{ST}$, are themselves summaries of the SFS.
Thus, the development of methods to learn demographic history from the SFS came
soon after the publication of whole-genome sequencing from multiple individuals
\cite{marth2004allele, williamson2005simultaneous}. More recent methodological
advances allow for increased sample sizes and numbers of populations, providing
a rich ecosystem of software for SFS-based demographic inference
\cite{gutenkunst2009inferring, excoffier2011fastsimcoal,
    gravel2011demographic, jouganous2017inferring, ragsdale2018genomic,
kamm2020efficiently, dilber2024faster}. We return to this in the final section:
Considerations and Caveats.

This chapter focuses on multi-population demographic inference from the SFS
using \moments \cite{jouganous2017inferring}. Briefly, \moments computes the
expected SFS by numerically solving a system of ordinary differential equations
closely related to the diffusion approximation \cite{gutenkunst2009inferring}.
The flexibility of the numerical methods in \moments allows for non-constant
population size histories including up to five populations at a time, with gene
flow. Selection can also be incorporated, allowing for recessive, additive or
dominant selection, or over- or underdominance at a locus. This differs from
most coalescent-based approaches (e.g.,
\cite{excoffier2011fastsimcoal,dewitt2021nonparametric,dilber2024faster}),
which are typically limited to neutrality.

Many evolutionary mechanisms are known to affect allele frequency dynamics and
the SFS. This presents a challenge for demographic inference, as the observed
SFS may be distorted by non-demographic processes. However, this also presents
an opportunity to learn about different evolutionary processes, if demographic
history can be controlled for. The SFS has been used to infer the distribution
of fitness effects of new mutations, primarily for nonsynonymous variation in
coding regions (e.g., \cite{eyre2006distribution, boyko2008assessing,
kim2017inference}); to scan for historical selective sweeps (e.g.,
\cite{kim2002detecting, nielsen2005genomic}); and to understand how selection
on quantitative traits impacts the genetic architecture underlying those traits
(e.g., \cite{patel2024conditional, ragsdale2024archaic}). The SFS has also been
used to infer relative mutation rates and how they have changed over time
\cite{dewitt2021nonparametric}. Each of these analyses typically requires
partitioning the data in some way -- such as by functional annotation, mutation
context, local recombination rate -- so that comparisons can be made across SFS
from different classes of mutations.

\subsection*{Obtaining and installing the software}

\moments is available through \texttt{pypi} as \texttt{moments-popgen} (using
\texttt{pip install moments-popgen}). Source code can be downloaded from
\url{https://github.com/MomentsLD/moments}, and  documentation is hosted at
\url{https://momentsld.github.io/moments/}. The detailed online examples
accompanying this chapter can be found at
\url{https://github.com/StatisticalPopulationGenomics-2ndEd/moments/}, which uses
\moments version 1.4 and \demes version 0.2.

\subsection*{Interfacing \moments with \demes}

Specifying demographic models requires defining populations (or ``demes'') and
their relationships via splits and gene flow. \demes provides a standardized
and accessible format for defining demographic models \cite{gower2022demes},
and has been adopted by widely used software for population genetics simulation
and inference (including \msprime \cite{baumdicker2022efficient}, \momi
\cite{dilber2024faster}, \fwdpy \cite{thornton2019polygenic}, and
\texttt{GADMA} \cite{noskova2023gadma2}). \demes encourages interoperability
and reuse of code across software, ease of model implementation and reduction
of errors.

Demographic models are specified in YAML format (see \cite{gower2022demes} and
the associated documentation). To illustrate, the two-population
split-with-migration model depicted in Figure~\ref{fig:im}A is specified as
below:

\singlespacing
\begin{minted}[gobble=4, frame=single,]{yaml}
    description: split-with-migration example model
    generation_time: 29
    time_units: years
    demes:
    - name: ancestral
      epochs:
      - {start_size: 15000, end_time: 300000}
      - {start_size: 25000, end_time: 75000}
    - name: popA
      ancestors: [ancestral]
      epochs:
      - {start_size: 30000}
    - name: popB
      ancestors: [ancestral]
      epochs:
      - {start_size: 1200, end_size: 14000}
    migrations:
    - demes: [popA, popB]
      rate: 5e-5
\end{minted}
\doublespacing

In \moments, a \demes-specified demographic model can be directly used to
compute expectations for either the SFS or multi-population LD statistics
\cite{ragsdale2019models, ragsdale2020unbiased}. For the SFS, we simply need
to load the demographic model using \demes and then specify the number of
haploid samples to draw and from which populations.
\begin{minted}[gobble=4, frame=single]{python}
    import demes, moments
    g = demes.load('model.yaml')
    samples = {'popA': 60, 'popB': 60}
    fs = moments.Demes.SFS(g, samples=samples)
\end{minted}
Here, we have not specified a mutation rate. The default behavior sets $u=1$,
so that $\theta = 4N_eu=4N_e$, where $N_e$ is the initial size of the ancestral
population. With a known mutation rate, the SFS will be properly scaled by
passing the mutation rate as a keyword argument: \texttt{moments.Demes.SFS(g,
samples=samples, u=u)}. This is equivalent to scaling the output SFS above by
multiplying by \texttt{u}, as \texttt{u*fs}.

\section*{Examples: inference in \moments using \demes}

Below, we show two examples of multi-population demographic inference using the
joint SFS. The first example simulates data with known demographic parameters,
which we attempt to reinfer using both the original model and simpler
misspecified models. In the second example, we infer a relatively simple
Neanderthal-human history using data from two human populations and a
high-coverage ancient sample from the Neanderthal lineage. For both of these
examples, we briefly describe data processing and inference setup, emphasizing
the high-level components of the analyses. For detailed information about each
example, including managing data, specifying models, performing inference,
computing confidence intervals and visualizing fits, we refer readers to the
accompanying GitHub repository
(\url{https://github.com/StatisticalPopulationGenomics-2ndEd/moments}), which
is maintained with up-to-date versioning.

In order to optimize parameters in a \demes model, we need a way to specify
which parameters should be fitted. For this, we use a separate YAML-formatted
file to define \texttt{parameters}, each of which points to desired value(s) in
the demographic model and for which lower and upper bounds may be set. This
file also specifies any relative \texttt{constraints} between pairs of
parameters, to ensure the optimization routine only explores valid model space.
Below, we have included a truncated parameters file -- the full example is
available on GitHub (URL above).

\singlespacing
\begin{minted}[gobble=4, frame=single]{YAML}
    parameters:
    - name: T0
      values:
      - demes:
          ancestral:
            epochs:
              0: end_time
      lower_bound: 0
      upper_bound: 2e5
    - name: T1
      values:
      - demes:
          ancestral:
            epochs:
              1: end_time
      lower_bound: 0
      upper_bound: 1e6
    - name: Ne
      values:
      - demes:
          ancestral:
            epochs:
              0: start_size
      lower_bound: 1e2
      upper_bound: 1e6
    ...
    constraints:
    - params: [T0, T1]
      constraint: greater_than
\end{minted}
\doublespacing

Taking the demographic model, parameters file, and observed SFS together, we
can perform inference using the \texttt{moments.Demes.Inference.optimize()}
function. The initial parameter guesses are given by the input demographic
model, which may be perturbed using a keyword argument. The likelihood surface
may be rugged or flat in some dimensions, and optimization routines sometimes
fail to converge to the global (or sometimes even local) optimum. Thus, we
often try many different initial guesses for parameter values, which can be
randomly assigned using the \texttt{perturb} keyword argument. Running multiple
iterations of optimization with different optimization functions is also
recommended to ensure convergence.

Any value in the demographic model that is not specified in the parameters file
will remain unchanged, and is therefore treated as a fixed parameter. Here, we
also assume we have an estimate for the total mutation rate, $U$.

\singlespacing
\begin{minted}[gobble=4, frame=single]{python}
    import moments

    data = moments.Spectrum.from_file('data.fs')
    U = 1e-8 * 5e8  # per-base mutation rate times the total length

    model_file = 'model.yaml'
    params_file = 'params.yaml'
    output = 'model.fit.yaml'

    ret = moments.Demes.Inference.optimize(
        model_file, params_file, data,
        perturb=1, uL=U, output=output
    )
\end{minted}
\doublespacing

Confidence intervals may also be computed using
\texttt{moments.Demes.Inference.uncerts()}, once a (locally) optimal model has
been found. The following will emit a tab-separated file containing parameter
names, best-fit values and estimated standard errors.

\singlespacing
\begin{minted}[gobble=4, frame=single]{python}
    log_file = 'log.txt'

    uncerts = moments.Demes.Inference.uncerts(
        output, params_file, data,
        uL=U, output=log_file
    )
\end{minted}
\doublespacing

Uncertainties may be computed using either the Fisher Information Matrix, which
tends to underestimate confidence intervals due to nonindependence of linked
sites, or using the Godambe Information Matrix, which uses a set of
bootstrapped datasets to correct for this nonindependence in the data
\cite{coffman2016computationally}. Example 1 in the associated GitHub
repository demonstrates this using a simulated dataset.

\subsection*{Example 1: Inferring parameters in a simulated
split-with-migration model}

Using the split-with-migration model defined above (illustrated in
Figure~\ref{fig:im}A), we simulated data for 30 diploid individuals from both
populations using \msprime \cite{baumdicker2022efficient}. These simulations
consided of 500 regions, each of length 1 Mb, with per-base recombination and
mutation rates of $10^{-8}$. The joint SFS was computed by using \tskit's
\texttt{ts.allele\_frequency\_spectrum()} function separately on each replicate
region and summing across regions to find the total SFS across 500 Mb of data
(Figure~\ref{fig:im}B--D).

We fitted three demographic models to the simulated data. In addition to
reinferring the parameters from the simulated model, we fitted two simpler 
models to the same data. These had fewer parameters, both omitting the size 
change deeper in time in the ancestral population. This is meant to crudely 
mimic the scenario that we often face, in which the true history is more 
complicated than the parameterized model. It also highlights biases in 
inferences that can arise when features of the true history are not included.

When fitting the two misspecified models that do not allow for size changes in
the ancestral population, this split time is inferred to be substantially
deeper in the past than the true split time. This effect, as demonstrated in
\cite{momigliano2021biases}, is due to the decreased coalescence rate within
the ancestral population between the time of the expansion and divergence of
the descendent populations. These two misspecified models
(Figure~\ref{fig:im}H,I) also provide a worse fit to the data, as seen by the
lower log-likelihoods and the large residuals between model predictions and
data (Figure~\ref{fig:im}E-G).

\subsection*{Example 2: Inferring Neanderthal-human demographic parameters}

We used publicly available data to infer a demographic model for two
modern human populations (a northern European and western African population) 
and a Neanderthal individual. As shown below, even with just a single sample 
from the Neanderthal lineage, we are able to recover key parameters such as the
Neanderthal-human divergence time and the fraction of ancestry contributed to 
modern Europeans by Neanderthals.

We obtained modern human genome sequences from a recent resequencing of the
1000 Genomes Project cohort \cite{byrska2022high}, and used the high-coverage
sequence of the Vindija Neanderthal genome \cite{prufer2017high}.  From the
1000 Genomes cohort we took 85 MSL (Mende from Sierre Leone) and 91 GBR
(British from England and Scotland) diploid sequences as our modern human
samples. We filtered out sites that lay outside the 1000 Genomes `strict' mask,
were not genotyped in four high-coverage archaic genome sequences, lacked
high-confidence ancestral state assignments, or fell within 10 kb of exonic or
promoter regions. This left us with approximately 960 Mb of well-characterized
and putatively neutrally-evolving sequence. We estimated the genome-wide SFS
using built-in \moments parsing functions. Details concerning data processing
(including our liftover of the Vindija sequence to a more recent genome build)
can be found in the linked GitHub repository.

The model that we wished to fit is relatively simple but still has many
parameters ($11$ in total), so we constructed it in stages by first fitting
marginal one- and two-population models. When we want to infer a complex
demographic history involving multiple populations, it is helpful to start
simple by modeling subsets of the full data and sequentially adding complexity.
This procedure gives us initial guesses for parameter values in more complex
models and may help suggest which features are well-supported by the data.
Here, we began by fitting one-population models with ancestral and recent size
changes for the MSL and GBR populations. In the MSL model, we allowed three
epochs of piecewise-constant effective size. We attributed to GBR two
constant-size epochs followed by a sharp contraction and exponential growth
beginning $\sim$60 kya to represent the out-of-Africa bottleneck and subsequent
rapid population expansion. Throughout, we used $u=1.5\times 10^{-8}$ as an
estimate of the mutation rate per generation and 29 years as the average
generation time. YAML files encoding the models and parameters in this section
can be found in the linked GitHub repository.

We next stitched these one-population models together and refitted the
resulting demographic model to get a two-population model that included
continuous symmetric migration between MSL and GBR and an ancestral size
expansion.  Following this, we added a Neanderthal branch and fitted the
Neanderthal-human divergence time, the Neanderthal effective size, and the
Neanderthal-GBR pulse proportion, leaving all other parameters fixed. We also
fitted the divergence time between the sampled Vindija Neanderthal lineage and
the Neanderthal lineage which admixed with non-African humans. Finding that
this parameter tended to converge to its lower bound (imposed by the estimated
time of death of the Vindija Neanderthal, ~55 kya), we fixed it to a value
supported by literature (90 kya, \cite{prufer2017high}) in a second round of
optimization. We also fixed the time of the Neanderthal-GBR pulse to 50 kya (at
the upper range of values inferrred by \cite{sumer2025earliest}).  Fixing
parameters for which good estimates are available in the literature is often
reasonable in the first stages of model optimization, and we expected this
value to be especially poorly constrained due to the small pulse proportion and
Neanderthal sample size. In the second round of optimization of the
three-population model, we refitted all parameters aside from the two times
mentioned above to refine the estimates.

After this refitting step, we were left with parameters in good agreement with
those inferred by prior authors, with a Neanderthal-human divergence time of
547 kya, MSL-GBR split time of 77 kya, and Neanderthal-GBR pulse proportion of
2.7 percent. We observe some large, systematic residuals in our best-fit model
(Figure~\ref{fig:humans}G-I), suggesting model misspecification that could be
improved through further exploration of model space.

We also fitted a nested model without the Neanderthal-GBR pulse, which had a
substantially lower likelihood than the pulse model (log-likelihood [LL]
$\approx -127,552$, against LL $\approx -25,103$ for the pulse model). We did
not perform a formal hypothesis test; the null distribution of the
likelihood-ratio test (LRT) statistic is overdispersed under a composite
likelihood, inflating the rate of false positives and preventing its use in
this setting. \moments does incorporate a function
(\texttt{moments.Godambe.LRT\_adjust()}) which computes a multiplicative
adjustment to the LRT, intended to restore the expected null distribution,
using first-order moment matching \cite{coffman2016computationally}. This
method requires bootstrapping over the data. Otherwise, we could conduct a
hypothesis test by fitting the nested and pulse models to a large number of
bootstrap samples and calculating the rate of rejection across them. This has
the benefit of producing a distribution of (nonindependent) LRT statistics, but
is time-consuming and computationally expensive.

\section*{Considerations and caveats}\label{sec:conclusions}

The site frequency spectrum is just one of many options for inferring
evolutionary parameters from genetic data. It is straightforward to compute
from data, and there are many methods for predicting the SFS, making it an
appealing framework for demographic inference. As shown here in two examples,
we have incorporated the \demes format for specifying demographic models into
\moments to easily compute the SFS and perform demographic inference, including
computing uncertainties.

\moments is limited in the size of the dataset that it can consider. Sample
sizes can be large (in the hundreds), especially when considering models with
one or few populations, but memory issues will develop when sample sizes are
extreme or as the number of populations grows. We are limited to a maximum of
five populations existing at any one time, including ancient samples. Other
SFS-based methods, such as \momi \cite{kamm2020efficiently, dilber2024faster},
can include additional populations, typically with smaller sample sizes.

When inferring models with multiple populations, especially four or more, both
the numbers of parameters and the space of plausible model topologies can grow
rapidly. It can therefore become challenging for the optimization routines to
find optima in parameter space, and overparameterization can become a problem.
One hint that models are overparameterized (or models misspecified) is runaway
behavior of parameter values, in which they converge to unreasonably small or
large values (or approach the imposed bounds). Starting with simple
parameterizations and sequentially adding complexity helps to alleviate this
issue. The challenge of large model space is more difficult to address, and
must be handled through building and testing a wide range of plausible model
topologies. While time-consuming, a broad exploration of model space,
especially when considering many populations, is crucial for evaluating
robustness of particular features of inferred history.

Finally, natural selection distorts allele frequencies not only at sites
directly under selection, but also in linked regions of the genome through
background selection. This can have a sizeable impact on the SFS and lead to
biases in demographic inference \cite{ewing2016consequences, johri2021impact}.
In gene-sparse species, such as primates, considering data from regions away
from functionally constrained elements is feasible, as we did in example 2
here. Alternatively, the genome may be partitioned by $B$-value (a measure of
the local strength of background selection, \cite{mcvicker2009widespread}) and
regions experiencing low levels of background selection may be used (for an
example, see \cite{medina2023demographic}). In some species with gene-dense
genomes, like \emph{Drosophila} species, this may be infeasible, and
interspersed putatively neutral variation must be used, such as synonymous
variation or short introns, which will still experience background selection.
The appropriate strategy will depend on the genome architecture of the species
and genomic resources available.

\bibliographystyle{unsrt}
\bibliography{paper}

\clearpage

\begin{figure}
    %\includegraphics{../example1/fig1.pdf}
    \caption{
        \textbf{Inference using simulated data.}
        (A) The true demographic model, including an expansion in the ancestral
        population and symmetric migration between the descendent populations.
        (B--D) The simulated data, showing both marginal and the joint SFS.
        (E--G) Residuals between the data SFS (in D) and the predicted SFS from
        the inferred demographic models in (H--J), resp. The two misspecified
        models (H and I) provide worse fits to the data and both overestimate
        the split time due to disallowing any size change in the ancestral
        population.
    }
    \label{fig:im}
\end{figure}

\begin{figure}
    %\includegraphics{../example2/fig2.pdf}
    \caption{
        \textbf{Inference using human and Neanderthal data.}
        (A) The best-fit model including the Vindija Neanderthal and two human
        populations.
        (B, C) Marginal SFS fits between the model and empirical data are shown
        for MSL and GBR.
        (D--I) Marginal two-way spectra predicted by the model and their residuals 
        against the data are plotted for the three pairs of populations MSL, GBR 
        (D and G), MSL, Vindija (E and H), and GBR, Vindija (F, I).
    }
    \label{fig:humans}
\end{figure}

\end{document}